A boundary-integral framework to simulate \\ viscous erosion of a porous medium

### 2018-04-10 docs/methods/erosion_methods.tex
NM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Flow-induced erosion deteriorates and reshapes solid material over a range of scales found in nature, from large geological formations~\cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015}, down to human-scale features and patterns~\cite{daerr2003erosion}. Though less visible, these same forces are working at the very smallest scales, slowly deteriorating the individual constituents of porous media (e.g.~solid, sand, or clay) or biological structures like plaque and biofilms~\cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such examples, this paper develops numerical methods to simulate fluid-mechanical erosion of multiple solid bodies in a Stokes flow---the most relevant regime for groundwater and the aforementioned bio-fluid applications.

Recent work in the high-Reynolds-number regime has highlighted the importance of a shape-flow feedback that occurs during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann} and the related processes of dissolution and melting~\cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}.  In these cases, the flow alters the morphology of the immersed structures, which in turn modifies the surrounding flow, and so on. Such processes of `sculpting with flow' \cite{ristroph2018sculpting} have been examined in the context of a single body \cite{ris-moo-chi-she-zha2012} and, more recently, multiple bodies whose shape evolution is coupled through the flow \cite{hewett2017evolution, moore2017riemann}.
From this viewpoint, the example of an eroding porous medium presents new opportunities and challenges. First, since the medium is composed of many individual constituents, multi-body interactions must be considered. Second, the long-range effects inherent to Stokes flow implies that the shape-evolution of bodies are even more strongly coupled than in any other regime.
% in particular close-contacts?

@@

\paragraph{Related Work}
In closely related work, Mitchell \& Spagnolie used a traction integral equation formulation to simulate viscous erosion of a single body in three dimensions \cite{mit-spa2016}. Their formulation was flexible enough to simulate a {\em mobile} body that erodes in various background flows (uniform flow, shear flow, etc.). An advantage of the traction formulation is that is directly provides the shear stress that is needed to evolve surfaces.  In the present work, we aim to simulate the erosion of a porous medium and further we choose to fix the bodies to their initial locations. Thus, instead than simulating a single, mobile body as in \cite{mit-spa2016}, we will simulate the erosion of many, immobile bodies. Furthermore, we will employ a standard double-layer formulation, as it allows us to easily calculate bulk quantities such as vorticity, pressure, and velocity in a post-processing step.

For these porous network simulations, it is essential to resolve the nonlocal effects between the different bodies.  BIEs naturally resolve these interactions, and they have been used extensively to simulate viscous fluids in complex geometries, including porous media flow~\cite{dea-qua-bir-jua2018, bar-mar-vee-zha2018}.  Common challenges of a BIE formulation include developing efficient preconditioners for the discretized system and numerical difficulties associated with nearly touching bodies. In this work, we consider geometries requiring $\mathcal{O}(100)$ GMRES iterations, so that preconditioners are not necessary. In future work, we will address mobile bodies for which near contacts may necessitate preconditoners~\cite{qua-bir2015a, qua-cou-dar2018, cou-pou-dar2017}. Similarly, near-singular schemes will not be necessary due to the fact that the spacing between receding bodies of fixed position always grows in time. Again, the future consideration of mobile bodies may necessitate near-singular schemes~\cite{qua-bir2014a, klo-bar-gre-one2013, bar-wu-vee2015, hel-oja2008a, bea-lai2001}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Flow-induced erosion deteriorates and reshapes solid material over a range of scales found in nature, from large-scale land formations~\cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015}, down to human-scale features and patterns~\cite{daerr2003erosion}. Though less visible, these same forces are working at the very smallest scales, slowly deteriorating the individual constituents of porous media (e.g.~solid, sand, or clay) or biological structures like plaque and biofilms~\cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such examples, this paper develops numerical methods to simulate fluid-mechanical erosion of multiple solid bodies in a Stokes flow---the most relevant regime for groundwater and the aforementioned bio-fluid applications.

Recent work in the high-Reynolds-number regime has highlighted the importance of a shape-flow feedback that occurs during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann} and the related processes of dissolution and melting~\cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}.  In these cases, the flow alters the morphology of the immersed structures, which in turn modifies the surrounding flow, and so on. Such processes of `sculpting with flow' \cite{ristroph2018sculpting} have been examined in the context of a single body \cite{ris-moo-chi-she-zha2012, moore2017riemann} and, more recently, multiple bodies whose shape evolution is coupled through the flow \cite{hewett2017evolution}.
From this viewpoint, the example of an eroding porous medium presents new opportunities and challenges. First, since the medium is composed of many individual constituents, multi-body interactions must be considered. Second, the long-range effects inherent to Stokes flow implies that the shape-evolution of bodies are even more strongly coupled than in any other regime.
% in particular close-contacts?

@@

\paragraph{Related Work}
In closely related work, Mitchell \& Spagnolie used a traction integral equation formulation to simulate viscous erosion of a single body in three dimensions \cite{mit-spa2016}. The flexibility of their formulation enabled simulations of a {\em mobile} body eroding in an {\em aribtrary} background flow (e.g.~uniform flow, shear flow, etc.). Another advantage of the traction formulation is that is directly provides the shear stress that is needed to evolve surfaces.  In the present work, we aim to simulate the erosion of a porous medium and, further, choose to fix the bodies to their initial locations. Thus, rather than a single, mobile body as considered in \cite{mit-spa2016}, we will simulate the erosion of many, immobile bodies. Although the traction formulation has several attractive features, we will instead employ a standard double-layer formulation, since it allows us to easily calculate bulk quantities such as vorticity, pressure, and velocity, that will facilitate our understanding of the erosion.

For the porous network simulations considered here, it is essential to resolve nonlocal effects between the different bodies.  BIEs naturally resolve these interactions, and they have been used extensively to simulate viscous fluids in complex geometries, including porous media flow~\cite{dea-qua-bir-jua2018, bar-mar-vee-zha2018}.  Common challenges of a BIE formulation include developing efficient preconditioners for the discretized system and numerical difficulties associated with nearly touching bodies. In this work, we consider geometries requiring $\mathcal{O}(100)$ GMRES iterations, so that preconditioners are not necessary. In future work, we will address mobile bodies for which near contacts may necessitate preconditoners~\cite{qua-bir2015a, qua-cou-dar2018, cou-pou-dar2017}. Similarly, near-singular schemes will not be necessary due to the fact that the spacing between receding bodies of fixed position always grows in time. Again, the future consideration of mobile bodies may necessitate near-singular schemes~\cite{qua-bir2014a, klo-bar-gre-one2013, bar-wu-vee2015, hel-oja2008a, bea-lai2001}.
